Library Parser

### 2024-12-12 studies/library_parser.qmd
fix(library_parser): update examples, remove outdated code

**Status:** in progress.

**TODO**: \
- Separate scripts and functions into separate .qmd files/directories. \
- Update `model_rate_emotion_values` nomenclature process for classification papers.\
- Replace rbind with custom bind function throughout

@@

```{r results = 'hide'}
library(stringr)
library(knitr, include.only = 'kable')
```

@@

```{r}
## AT LEAST 2 MISSING FIELDS
check_missing(2) -> missing_2

missing_2 |> dplyr::tibble()
```

```{r}
## AT LEAST 3 MISSING FIELDS
check_missing(3) |> dplyr::tibble()
```

# Outdated: Dealing with nested bibtex fields.

Some studies include multiple bibtex fields for substudies/evaluations of different feature sets. E.g., `MODEL_RATE_EMOTION_VALUES_A` and `MODEL_RATE_EMOTION_VALUES_B`. We need to apply special operations to separate these values for further analysis.

```{r eval = F, echo = F}
## function to parse output of multiple models nested within a single cell
parse_model_output = function(df)
{ # loop across rows
  # apply function to replace NAs with empty string
  sapply(1:nrow(df), function(x) {
    df$model_rate_emotion_values <- tidyr::replace_na(df$model_rate_emotion_values, '') # detect cells with rbind in them
    if(str_count(df$model_rate_emotion_values[x], "rbind") > 1) {
      # match first model (name between first underscore and rbind)
      model_1_name <- str_match(df$model_rate_emotion_values[x], 
                                "(?<=_).*?(?=rbind)")
      # match subsequent models (name between MODEL_RATE_EMOTION_VALUES_ and rbind)
      model_names <- str_extract_all(df$model_rate_emotion_values[x], 
                                     "(?<=MODEL_RATE_EMOTION_VALUES_).*?(?=rbind)")
      # combine names
      model_names <- unlist(c(model_1_name, model_names))
      # remove special characters
      model_names <- gsub("[^[:alnum:]]", "", model_names)
      # get corresponding values for each model
      model_results <- str_split(df$model_rate_emotion_values[x], 
                "MODEL_RATE_EMOTION_VALUES")
      # get rid of everything up to "rbind"
      model_results <- lapply(model_results, function(x) str_remove(x, 
                                                          '.+?(?=rbind)'))
      # now unlist the models and give them names to add later
      model_results <- unlist(model_results)
       model_results <- paste(model_names, '=', model_results)
      names(model_results) <- model_names
     # return(model_results)
    }
    else { # otherwise, if dealing with non-listed, just return the existing value
      model_results <- df$model_rate_emotion_values[x]
    }
    return(model_results)
  }
    )
  # see whether rbind appears multiple times

}

# get results from nested models:
model_outputs <- parse_model_output(meta_df)
# assign unique names
model_names <- lapply(model_outputs, names)

@@

## Data Frame Expansion

Next we want to copy the number of rows for each of the bibtex cells requiring special nesting

```{r eval = F, echo = F}
# count the numbers 
n_entries <- unlist(lapply(model_outputs, function(x) length(x)))
# for those returning 0, change value to 1
n_entries[n_entries == 0]<- 1
# replicate rows based on number of models reported
meta_df_unpacked <- meta_df[rep(seq_len(nrow(meta_df)), n_entries), ]
# now overwrite model_rate_emotion_values with the output of parse_model_output
meta_df_unpacked$model_rate_emotion_values <- unlist(model_outputs)

## Sanity check

```{r eval = F, echo = F}
# ex_substudy_1 <- eval(parse(text = meta_df_unpacked[7, 'model_rate_emotion_values']))
# ex_substudy_2 <- eval(parse(text = meta_df_unpacked[8, 'model_rate_emotion_values']))
```

## Print output of all studies:

```{r eval = F, echo = F}
sapply(1:nrow(meta_df_unpacked), function(i) {
  print(i) 
  print(eval(parse(text = meta_df_unpacked$model_rate_emotion_values[i])))
}
)
```

```{r eval = F, echo = F}
classification_studies = subset(meta_df_unpacked, paradigm == " classification  ")

sapply(1:nrow(classification_studies), function(i) {
  print(i) 
  print(eval(parse(text = classification_studies$model_rate_emotion_values[i])))
}
)
```

```{r eval = F, echo = F}
regression_studies = subset(meta_df_unpacked, paradigm == " regression  ")

sapply(1:nrow(regression_studies), function(i) {
  print(i) 
  print(eval(parse(text = regression_studies$model_rate_emotion_values[i])))
}
)
```

```{r}
# low-level function to extract relevant values from named array
model_result_2_df <- function(x) {
  # evaluate the expression
  x <- rlang::eval_tidy(rlang::parse_expr(x))
  # get name of fitted models from column names
  model_names <- colnames(x)
  # print(model_names)
  # split across model name for summary statistic (e.g., mean, sd, etc.)
  model_statistic <- str_split(model_names, '\\.')
  model_statistic <- unlist(lapply(model_statistic, function(x) x[2]))
  # get values
  model_values <- as.numeric(x)
  # names(model_values) <- 'score'
  print(model_values)
  # get additional attributes (feature.data.exp)
  model_attributes <- rownames(x)
  data.frame(model_attributes)
  # get additional model details
  model_attributes <- data.frame(do.call('rbind', str_split(model_attributes, '\\.')))
  names(model_attributes) <- c('library_id', 'model_id', 'feature_id', 'data_id', 'experiment_id')
  print(names(model_attributes))
  # split across '_' to 
  model_measures <- data.frame(do.call('rbind', stringr::str_split(colnames(x), '_')))
  print(model_measures)
  names(model_measures) <- c('dimension', 'measure')
  model_measures_split <- stringr::str_split(model_measures[,'measure'], '\\.')
  model_measures[,'measure'] <- unlist(lapply(model_measures_split, 
                                              function(x) x[1]))
  return(data.frame(
             model_attributes, 
             model_measures, 
             values = model_values,
             statistic = model_statistic))
}

# high level function to apply model_result_2_df to multiple studies
get_study_summaries <- function(df) {
  do.call(rbind,
          lapply(df$model_rate_emotion_values, 
                 FUN = function(x) {
                   study_id <- unique(df$citekey[which(df$model_rate_emotion_values == x)])
                   model_results <- model_result_2_df(x)
                   return(cbind(study_id, model_results))
                   }
                 )
          )
}

# Parsing Regression Studies

The following functions provide a working example for converting the r code contained in `model_rate_emotion_values` to a data frame for further analysis (currently working for regression studies).

### Template

Encoded values should use the following nomenclature:

```         
library = rbind(library.model.features.data.exp = c(dim_measure.summary = 0.5, dim_measure.summary = 0.1), library.model.features.data.exp = c(dim_measure.summary = 0.2, dim_measure.summary = 0.5))

# e.g.

rbind(librosa.randomforest.mfcc.msd.1 = c(valence_r2.mean = 0.5, arousal_r2.mean = 0.1), 
librosa.randomforest.temporal.msd.1 = c(valence_r2.mean = 0.2, arousal_r2.mean = 0.5,
valence_r2.sd = 0.1, arousal_r2.sd = 0.1))
```

### Working example

```{r}
this_study <- data.frame(citekey = c('testStudy2023ab','testStudy2023ab', 'testStudy2020of'),
  model_rate_emotion_values = c("rbind('marsyas.random forest.mfcc.msd.1' = c(valence_r2.mean = 0.9, arousal_r2.mean = 0.5,
                                  valence_r2.sd = 0.4, arousal_r2.sd = 0.3,
  valence_mse.mean = 0.3, arousal_mse.mean = 9.1),
  'random forest.mfcc.deam.1' = c(valence_r2.mean = 0.9, arousal_r2.mean = 0.5))",
  "rbind('librosa.linear regression.mfcc.ismir.1' = c(valence_r2.mean = 0.1, arousal_r2.mean = 0.3, valence_r2.sd = 0.2, arousal_r2.sd = 0.4,
  euclid_euclid.mean = 0.32, euclid_euclid.sd = 0.1))",
  "rbind('mirtoolbox.neural net.pitch.filmMusic.2' = c(valence_r2.mean = 0.1, arousal_r2.mean = 0.3))")
)
```

### Summary data frame

```{r}
dplyr::tibble(get_study_summaries(this_study))
```

>>>>

@@

```{r results = 'hide'}
library(stringr)
library(knitr, include.only = 'kable')
source(paste0(here::here(), "/R/format-study-results.R"))
source(paste0(here::here(), "/R/parse-model-output.R"))

@@

```{r}
## AT LEAST 2 MISSING FIELDS
check_missing(2) |> dplyr::tibble()
## AT LEAST 3 MISSING FIELDS
check_missing(3) |> dplyr::tibble()
```

# Parsing Studies

The following functions provide a working example for converting the *R* code contained in `model_rate_emotion_values` to a data frame for further analysis. `get_study_summaries` is a simple function to demonstrate how the preprocessing routines work in the following examples

```{r}
# high level function to apply parse_model_output to multiple studies
get_study_summaries <- function(df) {
do.call(rbind,
  lapply(df$model_rate_emotion_values, 
         FUN = function(x) {
           study_id <- unique(df$citekey[which(df$model_rate_emotion_values == x)])
           model_results <- parse_model_output(x)
           return(cbind(study_id, model_results))
           }
         )
  )
}

@@

## Regression studies

Encoded values should use the following nomenclature. In this synthetic example, dim_measure.summary2 exists for the study in the second row, but not the first, so an NA is encoded for summary2 in row 1.

```{r warning=FALSE}
bind_field(
  'library.model.features.data.exp1' = c(dim_measure1.summary = 0.5, dim_measure2.summary = 0.1), 
  'library.model.features.data.exp2' = c(dim_measure1.summary = 0.2, dim_measure2.summary = 0.5, dim_measure.summary2 = 0.3)
)
```

### Example 2

This example is more extensive, and shows how this procedure can work with multiple studies:

```{r}
regression_example <- data.frame(citekey = c('testStudy2023aa','testStudy2023bb', 'testStudy2020cc'),
  model_rate_emotion_values = c("bind_field('marsyas.random forest.mfcc.msd.1' = c(valence_r2.mean = 0.9, arousal_r2.mean = 0.5,
 valence_r2.sd = 0.4, arousal_r2.sd = 0.3),
  'essentia.random forest.mfcc.deam.1' = c(valence_r2.mean = 0.9, arousal_r2.mean = 0.5))",
  "bind_field('librosa.linear regression.mfcc.ismir.1' = c(valence_r2.mean = 0.1, arousal_r2.mean = 0.3, euclid_euclid.mean = 0.32))",
  "bind_field('mirtoolbox.neural net.pitch.filmMusic.2' = c(valence_r2.mean = 0.1, arousal_r2.mean = 0.3))")
)

dplyr::tibble(get_study_summaries(regression_example)) |>
  knitr::kable()
```

## Classification studies

For classification studies, we can extract results from a matrix-like representation.

`unflatten` is a convenience function for preparing a confusion matrix, and `summarize_matrix` calls on the *caret* package to compute summary statistics for the matrix.

```{r warning = FALSE}
bind_field(
  lapply(
    list(
      'library.model.features.data.exp' = unflatten(
      "Class_A" = 0.8, "Class_B" = 0.2,
      0.2, 0.8
      ),
      'library.model.features.data.exp' = unflatten(
      "Class_A" = 0.4, "Class_B" = 0.6,
      0.6, 0.4
      )
    ),
    summarize_matrix
  )
)
```

```{r warning = FALSE}

classification_example <- data.frame(
  citekey = "classificationStudy2020aa",
  model_rate_emotion_values = 
  "bind_field(
    lapply(
      list(
        'marsyas.smo unambiguous.mixed.new.1' = unflatten(
          'A' = 52.6, 'B' = 17.1, 'C' = 0.0,
          12, 65.2, 7.6, 
          1.1,9.9,73.6),
        'marsyas.smo full.mixed.new.1' = unflatten(
          'C1' = 56.0, 'C2' = 19, 'C3' = 0.0, 'C4' = 2, 
          11.3,58.9,13.7,15,
          0,13.2,68.6, 3,
          15,1,4, 6)
      ),
      summarize_matrix
    )
  )"
)

dplyr::tibble(get_study_summaries(classification_example)) |>
  knitr::kable()
```
